---
title: "BotanizeR functions"
author: "Patrick Weigelt, Pierre Denelle"
date: "`r Sys.Date()`"
output: 
  html_vignette:
    toc: true
    toc_depth: 2
    number_sections: true
vignette: >
 %\VignetteIndexEntry{BotanizeR functions}
 %\VignetteEncoding{UTF-8} 
 %\VignetteEngine{knitr::rmarkdown}
editor_options: 
 chunk_output_type: console
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 5)
# Packages --------------------------------------------------------------------
suppressPackageStartupMessages({
  suppressWarnings({
    library(BotanizeR)
    library(sf)
  })
})

options(tinytex.verbose = TRUE)

#devtools::load_all()
```


# Overview  

\newpage

BotanizeR is an R-package with Shiny app designed to help botany students and 
other people interested in plants to learn and distinguish plant diagnostic 
characters, memorize plant species, and train their identification skills. 
Students can browse a species list, images, plant characteristics, habitat 
descriptions and distribution information or play a quiz where images of a 
random species are shown and students have to guess the correct species. 
BotanizeR is highly flexible allowing lecturers to define their own species 
lists and to provide images and further useful information, or choose from 
linked online content from selected botanical online resources.
A Quiz and Species overview pages can be made available via the package’s Shiny 
application or played within R directly. 

# Installation
```{r Installation, eval=FALSE}
library(devtools)
install_github("https://github.com/patrickweigelt/BotanizeR")

library(BotanizeR)
```



# Functions

The BotanizeR R package includes functions to collect species level information 
from defined reources, to play the BotanizeR quiz, to prepare content for the 
quiz (batch resizing images, creating a local species list) and for running the 
BotanizeR Shiny app.  


## BotanizeR_species

`BotanizeR_species` is an example dataset comprising 
a species list containing the relevant information to run 
`BotanizeR_quiz` or the BotanizeR Shiny app for 5852 
vascular plant species from [FloraWeb](https://www.floraweb.de) 
and the [Online Atlas of the British Irish flora](https://www.brc.ac.uk/
plantatlas/).

```{r BotanizeR_species}
data(BotanizeR_species)
BotanizeR_species[c(1:4), 
                  c("NAMNR", # FloraWeb ID
                    "TAXONNAME","SPECIES","GENUS", # Names to display and guess
                    "UK_Ireland", "Germany", # Columns to make subsets
                    "COUNT", "SCORE", "ATTEMPTS", "INCLUDE", # Quiz controls            
                    "ownhint_English_name", "ownhint_Description", # Own hints
                    "imagelink_1", "imagelink_2")] # Own image links
```

## BotanizeR_getlocallist

`BotanizeR_getlocallist()` uses longitude and latitude coordinates and a defined 
buffer radius to retrieve a plant species list from 
[GBIF](https://www.gbif.org/) and formats it for use in 
`BotanizeR_quiz` or the BotanizeR Shiny app. If a `backbone_list` 
is supplied, it will be subset for the species found in 
[GBIF](https://www.gbif.org/) and returned.

```{r BotanizeR_getlocallist}
# Subset the species list for species occurring in Cairngorms National Park
BotanizeR_Cairngorms <- 
  BotanizeR_getlocallist(long = -3.5966, lat = 57.0877, 
                         radius = 0.5, taxonKey = 7707728, 
                         limit = 500, backbone_list = BotanizeR_species)
nrow(BotanizeR_Cairngorms)


# Create a new species list for the Island of Bornholm, Denmark
BotanizeR_Bornholm <- 
  BotanizeR_getlocallist(long = 14.9184, lat = 55.1273, 
                         radius = 0.5, taxonKey = 7707728, 
                         limit = 500, backbone_list = NA)
nrow(BotanizeR_Bornholm)
```

## BotanizeR_imageresize

`BotanizeR_imageresize()` batch resizes all jpg and png images found in 
defined folders for plotting in `BotanizeR_quiz()` or the BotanizeR Shiny app.  
You define the new width of the images and optionally a maximum height which 
shall not be exceeded. Aspect ratios are maintained. `BotanizeR_quiz()` and the 
BotanizeR Shiny app can both resize images on the fly, however, to reduce 
traffic and loading time, it makes sense to reduce the size of large images in 
advance using `BotanizeR_imageresize()`.

```{r BotanizeR_imageresize, eval=FALSE}
## Not run: 
BotanizeR_imageresize(
  image_folders = c("inst/WWW/images_1",  # define 1 ore more folders
                    "inst/WWW/images_2"), # with images to resize 
                      image_width = 560, max_height = 700, # new maximum extent
                      quality = 1, int_type = 5) # quality and interpolation 
                                                 # type as in imager::resize()

## End(Not run)
```

## BotanizeR_collect

`BotanizeR_collect()` Collects information from online resources like 
[FloraWeb](https://www.floraweb.de) (images, map and species descriptions) 
and/or the [Online Atlas of the British Irish flora](https://www.brc.ac.uk/
plantatlas/) as well as from user defined image folders and columns in the 
`species_list` data.frame and formats them for showing them as hints in 
`BotanizeR_quiz()` or the BotanizeR Shiny app. This function does not have to 
be run standalone outside the quiz, but running it exemplifies how 
`BotanizeR_quiz()` and the BotanizeR Shiny app retrieve the information to be 
shown and how own information needs to be supplied via image folders or colunms 
of the `species_list` data.frame.


```{r BotanizeR_collect}
# Select Acer campestre as example species
species_row = BotanizeR_species[which(BotanizeR_species$SPECIES ==
                                        "Acer campestre"),]


# Load hints: only UK plant atlas image(s) + hints + map
hints <- BotanizeR_collect(species_row, image_ukplantatlas = TRUE, 
                           hints_ukplantatlas = c("mapuk", "familyuk", 
                                                 "ecology", "statusuk", 
                                                 "trends", "perennation",
                                                 "lifeform", "woodiness", 
                                                 "clonality"))

par(mar = rep(0.5, 4), oma = rep(0, 4))
plot(hints$image[[2]], axes = FALSE)

plot(hints$mapuk)

hints$statusuk
hints$clonality


# Load hints: only FloraWeb image(s) + hints + map
hints <- BotanizeR_collect(species_row, image_floraweb = TRUE, 
                           hints_floraweb = c("map", "description", 
                                              "status", "habitat", 
                                              "family", "German name"))

par(mar = rep(0.5, 4), oma = rep(0, 4))
plot(hints$image[[1]], axes = FALSE)

par(oma = c(0, 0, 0, 10.5))
plot(hints$map[[1]], pal = hints$map[[2]], key.pos = 4, main = "")

hints$family


# Load hints: only images from custom image links + custom hints
hints <- BotanizeR_collect(species_row, image_floraweb = FALSE,
                           hints_custom = c("ownhint_English_name", 
                                            "ownhint_Description", 
                                            "ownhint_Distribution"),
                           imagelinks_custom = c("imagelink_1", 
                                                 "imagelink_2"))
par(mar = rep(0.5, 4), oma = rep(0, 4))
plot(hints$image[[2]], axes = FALSE)

hints$ownhint_English_name


# Load hints: only returning image links instead of actual images
hints <- BotanizeR_collect(species_row, image_floraweb = TRUE,
                           image_ukplantatlas = TRUE, 
                           imagelinks_custom = c("imagelink_1", 
                                                 "imagelink_2"),
                           hints_ukplantatlas = c("mapuk", "familyuk"),
                           only_links = TRUE)
hints$images[[1]]

# To load images from your local computer, specify an image folder with
# pictures included. File names need to include the species names.

## Not run: 
hints <- BotanizeR_collect(species_row, 
                           image_folders = c("images/Asteraceae_Britain",
                                             "images/Trees"))
## End(Not run)
```


## BotanizeR_quiz

`BotanizeR_quiz()` runs the console version of the BotanizeR quiz. It navigates 
randomly through a `species_list`, shows defined pictures and descriptions and 
other hints from `BotanizeR_collect()` and let's you guess the species 
name. The Number of tries and attempts are used to calculate scores that will 
be used to update probabilities for the random sampling of subsequent species. 
The better you know a species the less likely it will be shown again.  
<br>
After running the main function BotanizeR_quiz() you need to click into the 
console to type the species names in there. If you have no clou, press _enter_ 
and the next hint will appear (you have `max_attempts` tries). If you type the 
name slightly wrong the function will tell you. If the genus is correct it will 
also tell you. It shows several photos if available. If you want to skip a 
species write 'skip'. If you want to cancel the quiz write 'exit'. Don't hit 
_Esc_ if you want to save your progress.   

An example for species from Sussex, UK based on the 
[Online Atlas of the British Irish flora](https://www.brc.ac.uk/plantatlas/):  

```{r BotanizeR_quiz, eval=FALSE}

# Subset for about 700 species known from Sussex, UK
BotanizeR_Sussex <- 
  BotanizeR_species[which(BotanizeR_species$UK_Ireland_Sussex==1), ]


# Type in species name, or press enter for next hint or type 'skip' and
# press enter for next species or type 'exit' to end quiz and save results
BotanizeR_UK_practiced <- 
  BotanizeR_quiz(species_list = BotanizeR_example,
                 image_floraweb = FALSE, image_ukplantatlas = TRUE, 
                 hints_ukplantatlas =  c('mapuk', 'familyuk', 'ecology', 
                                         'statusuk', 'trends', 'perennation',
                                         'lifeform', 'woodiness', 
                                         'clonality'), 
                 hints_floraweb = NULL, 
                 case_sensitive = FALSE)

```

An example for species from Göttingen, Germany based on 
[FloraWeb](https://www.floraweb.de):  

```{r BotanizeR_quiz_Germany, eval=FALSE}
# Subset for about 300 species students in Goettingen, Germany, learn
BotanizeR_Germany <- 
  BotanizeR_species[which(BotanizeR_species$Germany_summer==1 | 
                            BotanizeR_species$Germany_BioDiv==1), ]


# Type in species name, or press enter for next hint or type "skip" and
# press enter for next species or type "exit" to end quiz and save results
BotanizeR_Germany_practiced <- 
  BotanizeR_quiz(species_list = BotanizeR_Germany, 
                 hints_floraweb = c("description", "status",
                                    "habitat","family","German name"), 
                 case_sensitive = FALSE)


# If you want to include distribution maps as hints add "map" to hints;
# This increases the download times a bit
BotanizeR_Germany_practiced <- 
  BotanizeR_quiz(species_list = BotanizeR_Germany,
                 hints_floraweb = c("map", "description", "status", 
                                    "habitat", "family", "German name"), 
                 case_sensitive = FALSE)

```

If you want to keep track of your progress, you can save the species list with 
updated scores locally and load it again in the next session:

```{r BotanizeR_quiz_saving, eval=FALSE}

## Not run: 

# Saving results from above
write.csv(BotanizeR_Germany_practiced, 
          "BotanizeR_Germany_practiced.csv", row.names = FALSE)

# Loading previously saved pratcicing results
BotanizeR_Germany_practiced <- read.csv("BotanizeR_Germany_practiced.csv")

# Practicing
BotanizeR_Germany_practiced <- 
  BotanizeR_quiz(species_list = BotanizeR_Germany_practiced, 
                 hints_floraweb = c("description", "status",
                                    "habitat","family","German name"), 
                 case_sensitive = FALSE)

# Saving results again
write.csv(BotanizeR_Germany_practiced, "BotanizeR_Germany_practiced.csv", 
          row.names = FALSE)

## End(Not run)```
```


You can add own species descriptions or other hints as well is link to 
additional online images into additional columns of the species list 
data.frame. Alternatively you can specify a folder with images on your local 
computer. Image file names need to include species names.

```{r custom, eval=FALSE}
# example for three species with custom hints and images from their Wikipedia page
BotanizeR_species_custom <-  BotanizeR_species[
  which(BotanizeR_Germany$species %in% c("Acer campestre", 
                                         "Erica carnea",
                                         "Melampyrum nemorosum")),]

custom_species_trained <- 
  BotanizeR_quiz(species_list = BotanizeR_species_custom, 
                 image_floraweb = TRUE, hints_floraweb = NULL, 
                 hints_custom = c("ownhint_Description","ownhint_Distribution"),
                 imagelinks_custom = c("imagelink_1", "imagelink_2"), 
                 image_folders = NULL, case_sensitive = FALSE)

```


## BotanizeR_shiny

The most convenient way to play the BotanizeR quiz and to study selected plant 
species based on images and descriptions is the BotanizeR Shiny app. When the 
BotanizeR R package is installed, you can start the Shiny application using the 
function `BotanizeR_shiny()`.  

```{r BotanizeR_shiny, eval=FALSE}
BotanizeR_shiny()
```

Alternatively, you can clone the BotanizeR GitHub 
repository (https://github.com/patrickweigelt/BotanizeR) and run the shiny app 
from RStudio. This option offers most flexibility for modifying the app. For 
teaching purposes the Shiny app would ideally be set up on a shiny server which 
students can visit online. A comprehensive example instance of BotanizeR 
showcasing all functionality can be found at 
https://gift.uni-goettingen.de/shiny/BotanizeR/ 

Two more tutorials, one on how to use the Shiny app and one on how to set it up 
can be found here:  

- **[how to use the BotanizeR Shiny app](https://patrickweigelt.github.io/BotanizeR/articles/BotanizeR_Shiny.html)**  
- **[how to set up and customize the Shiny app](https://patrickweigelt.github.io/BotanizeR/articles/BotanizeR_config.html)**  



# References

Bundesamt für Naturschutz (BfN): 
[http://www.floraweb.de](http://www.floraweb.de)  
Image authors: 
[https://www.floraweb.de/ueberfloraweb/bildautoren.html](https://www.floraweb.de/ueberfloraweb/bildautoren.html)

Online Atlas of the British and Irish flora: 
[https://www.brc.ac.uk/plantatlas/](https://www.brc.ac.uk/plantatlas/)  
Image authors: 
[https://www.brc.ac.uk/plantatlas/content/photos](https://www.brc.ac.uk/plantatlas/content/photos)  



