---
title: "BotanizeR config"
author: "Patrick Weigelt, Pierre Denelle"
date: "`r Sys.Date()`"
output: 
  html_vignette:
    toc: true
    toc_depth: 2
    number_sections: true
vignette: >
 %\VignetteIndexEntry{BotanizeR config}
 %\VignetteEncoding{UTF-8} 
 %\VignetteEngine{knitr::rmarkdown}
editor_options: 
 chunk_output_type: console
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 5)
# Packages --------------------------------------------------------------------
suppressPackageStartupMessages({
  suppressWarnings({
    library(BotanizeR)
  })
})

options(tinytex.verbose = TRUE)
```


# Overview  

\newpage

BotanizeR is an R-package with Shiny app designed to help botany students and 
other people interested in plants to learn and distinguish plant diagnostic 
characters, memorize plant species, and train their identification skills. 
Students can browse a species list, images, plant characteristics, habitat 
descriptions and distribution information or play a quiz where images of a 
random species are shown and students have to guess the correct species.

Lecturers can flexibly define their own species lists and provide images and 
further useful information, or choose from linked online content from selected 
botanical online resources to show in BotanizeR. In this Tutorial, we explain 
how to create own content to present in BotanizeR and how to customize the 
Shiny app.


# Shiny




# Species List

A new species list for BotanizeR needs to be a `data.frame` containing at least 
the following columns: *NAMNR*, *TAXONNAME*, *SPECIES* and *GENUS*. The *SPECIES* 
column includes the species name (genus and epithet; character or factor) to be 
guessed and looked up in the online resources or image folders. The *TAXONNAME* 
column includes the full species name including additional information like, 
for example the authority (character or factor). The *GENUS* column includes 
the corresponding genus name (character or factor). *NAMNR* contains the ID 
(numeric) of the species used by [FloraWeb](https://www.floraweb.de). In 
case FloraWeb content is not retrieved, this may be NA.

In addition, the species list  should include the numeric columns *COUNT*, 
*SCORE* and *ATTEMPTS* where the number of times a species was 
practiced, how often it was guessed correctly and how many attempts were needed 
will be recorded. At the start these columns include zeros only. In addition, 
the column *INCLUDE* indicates whether a species shall be included in the app 
or not. Here, the default value is 1.

In addition, the species list may contain columns to filter by, as 
well as columns including own descriptive hints (named like 
*ownhint_Description*, *ownhint_Distribution*, etc.) and links (URLs) to images 
(*imagelink_1*, *imagelink_2* etc.). These columns can be filled by the 
lecturer in advance but can also be modified and expanded by the student 
after downloading the list.

The R package includes an example species list 
containing the relevant information to run `BotanizeR_quiz` or the BotanizeR 
Shiny app for 5852 vascular plant species from 
[FloraWeb](https://www.floraweb.de) and the 
[Online Atlas of the British Irish flora](https://www.brc.ac.uk/plantatlas/).


```{r Installation, eval=FALSE}
library(BotanizeR)
```
```{r BotanizeR_species}
data(BotanizeR_species)
BotanizeR_species[c(13:15), 
                  c("NAMNR", # FloraWeb ID
                    "TAXONNAME","SPECIES","GENUS", # Names to display and guess
                    "UK_Ireland", "Germany", # Columns to make subsets
                    "COUNT", "SCORE", "ATTEMPTS", "INCLUDE", # Quiz controls            
                    "ownhint_English_name", "ownhint_Description", # Own hints
                    "imagelink_1", "imagelink_2")] # Own image links
```


# Images

Images can be supplied like indicated above by including image URLs into 
dedicated columns of the species list, or by supplying local folders with 
jpg or png images. When running the Shiny app locally (on the students 
computer) these folders need to be on the local computer. If the Shiny app is 
running on a server, these folders can be supplied by the lecturer on the 
server (see below).

In addition, Images can be sources from selected online resources currently
incorporated into BotanizeR: [FloraWeb](https://www.floraweb.de) and the 
[Online Atlas of the British Irish flora](https://www.brc.ac.uk/plantatlas/).
All images, from own image links, image folders, or online resources are 
collected for use in BotanizeR by the `BotanizeR_collect()` function which is 
called within the Shiny app and `BotanizeR_quiz()`:

```{r BotanizeR_collect, fig.width = 7, fig.height = 4}
# Select Acer campestre as example species
species_row = BotanizeR_species[which(BotanizeR_species$SPECIES ==
                                        "Acer campestre"),]


# Load images: UK plant atlas, FloraWeb and own image links
hints <- BotanizeR_collect(species_row, image_floraweb = TRUE, 
                           image_ukplantatlas = TRUE, 
                           imagelinks_custom = c("imagelink_1","imagelink_2"))

par(mfrow=c(2,3), mar = rep(0.2, 4), oma = rep(0, 4))
plot(hints$image[[1]], axes = FALSE)
plot(hints$image[[2]], axes = FALSE)
plot(hints$image[[3]], axes = FALSE)
plot(hints$image[[4]], axes = FALSE)
plot(hints$image[[5]], axes = FALSE)
plot(hints$image[[6]], axes = FALSE)


# To load images from your local computer, specify an image folder with
# pictures included.

## Not run: 
hints <- BotanizeR_collect(species_row, 
                           image_folders = c("images/Asteraceae_Britain",
                                             "images/Trees"))
## End(Not run)
```


# Descriptive hints

Descriptive hints are drawn from the ownhint columns of the species list 
data.frame as indicated above or from selected online resources currently
incorporated into BotanizeR: [FloraWeb](https://www.floraweb.de) and the 
[Online Atlas of the British Irish flora](https://www.brc.ac.uk/plantatlas/).
Again this information is collected by `BotanizeR_collect()` inside
the Shiny app and `BotanizeR_quiz()`:

```{r BotanizeR_collect_hints}

# Load hints: UK plant atlas, FloraWeb and own hints 
hints <- BotanizeR_collect(species_row,
                           hints_floraweb = c("description", 
                                              "status", "habitat", 
                                              "family", "German name"),
                           hints_ukplantatlas = c("familyuk", 
                                                  "ecology", "statusuk", 
                                                  "trends", "perennation",
                                                  "lifeform", "woodiness", 
                                                  "clonality"),
                           hints_custom = c("ownhint_English_name", 
                                            "ownhint_Description", 
                                            "ownhint_Distribution"))
hints$statusuk
hints$clonality
hints$lifeform
```


# BotanizeR Shiny app

## Getting started

The most convenient way to play the BotanizeR quiz and to study selected plant 
species based on images and descriptions is the BotanizeR Shiny app. When the 
BotanizeR R package is installed, the Shiny application can be started locally 
using the function `BotanizeR_shiny()`. Alternatively, one can clone the 
BotanizeR GitHub repository (https://github.com/patrickweigelt/BotanizeR) and 
run the shiny app from RStudio. This option offers most flexibility for 
modifying the app as all code can easily be changed. For teaching purposes the 
Shiny app would ideally be set up on a shiny server which students can visit 
online. 

A useful tutorial for setting up a shiny server can be found here:
https://www.digitalocean.com/community/tutorials/how-to-set-up-shiny-server-on-ubuntu-16-04

A comprehensive example instance of BotanizeR showcasing all functionality can 
be found at https://gift.uni-goettingen.de/shiny/BotanizeR/ 



## Setup tab



Bundesamt für Naturschutz (BfN): 
[http://www.floraweb.de](http://www.floraweb.de)  
Image authors: 
[https://www.floraweb.de/ueberfloraweb/bildautoren.html](https://www.floraweb.de/ueberfloraweb/bildautoren.html)

Online Atlas of the British and Irish flora: 
[https://www.brc.ac.uk/plantatlas/](https://www.brc.ac.uk/plantatlas/)  
Image authors: 
[https://www.brc.ac.uk/plantatlas/content/photos](https://www.brc.ac.uk/plantatlas/content/photos)  


