---
title: "BotanizeR tutorial"
author: "Patrick Weigelt"
date: "`r Sys.Date()`"
output: 
  html_vignette:
    toc: true
    toc_depth: 2
    number_sections: true
vignette: >
 %\VignetteIndexEntry{Tutorial for BotanizeR R package}
 %\VignetteEncoding{UTF-8} 
 %\VignetteEngine{knitr::rmarkdown}
editor_options: 
 chunk_output_type: console
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 8)
# Packages --------------------------------------------------------------------
suppressPackageStartupMessages({
  suppressWarnings({
    library(dplyr)
  })
})

options(tinytex.verbose = TRUE)

#devtools::load_all()
```


# Overview

BotanizeR.

\newpage

# Main function
## BotanizeR_quiz
After running the main function BotanizeR_quiz() you need to click into the console to type the species names in there. If you have no clou, press enter and the next hint will appear (you have ten tries). If you type the name slightly wrong the function will tell you. If the genus is correct it will also tell you. It shows several photos if available. If you want to skip a species write "skip". If you want to cancel the quiz write "exit". Don't hit Esc if you want to save your progress. The function counts your attempts and successes and uses this as sampling probabilities for choosing the species randomly. The better you know a species the less likely it will be shown again.

```{r example, eval=FALSE}
library(devtools)
install_github("https://github.com/patrickweigelt/BotanizeR")

library(BotanizeR)

# Species list for Germany with IDs from floraweb.de
data(floraweb_species)

# subset for about 300 species students in Göttingen learn
floraweb_species <- floraweb_species[which(floraweb_species$SUMMER==1 | floraweb_species$BioDiv2005==1),]

# Type in species name, or press enter for next hint or type "skip" and press 
# enter for next species or type "exit" to end quiz and save results
floraweb_species_trained <- BotanizeR_quiz(species_list = floraweb_species, 
                                           hints = c("description","status",
                                                     "habitat","family","German name"),
                                           case_sensitive = FALSE)
```


If you want to keep track of your progress, you can save the species list with updated scores locally and load it in the next session
```{r practicing, eval=FALSE}
# Initial saving
write.csv(floraweb_species_trained, "floraweb_species_trained.csv", row.names = FALSE)

# Load species list
floraweb_species_trained <- read.csv("floraweb_species_trained.csv")

# Practice
floraweb_species_trained <- BotanizeR_quiz(species_list = floraweb_species_trained,
                                           hints = c("description","status","habitat",
                                          "family","German name"), case_sensitive = FALSE)

# Save species list
write.csv(floraweb_species_trained, "floraweb_species_trained.csv", row.names = FALSE)
```



# References

Bundesamt für Naturschutz (BfN) [http://www.floraweb.de](http://www.floraweb.de)  

Image authors: [https://www.floraweb.de/ueberfloraweb/bildautoren.html](https://www.floraweb.de/ueberfloraweb/bildautoren.html)



